---
title: "Introduction to CodelistGenerator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction_to_CodelistGenerator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Getting the OMOP CDM vocabularies
To use CodelistGenerator we need to first obtain the OMOP CDM vocabularies from https://athena.ohdsi.org. Once these are downloaded, you can load in the relevant files like so:

```{r import vocabularies,  message=FALSE, warning=FALSE}
library(readr)
vocab.folder<-Sys.getenv("omop_cdm_vocab_path") # path to directory of unzipped files
concept<-read_delim(paste0(vocab.folder,"/CONCEPT.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
concept_relationship<-read_delim(paste0(vocab.folder,"/CONCEPT_RELATIONSHIP.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE) 
concept_ancestor<-read_delim(paste0(vocab.folder,"/CONCEPT_ANCESTOR.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
concept_synonym<-read_delim(paste0(vocab.folder,"/CONCEPT_SYNONYM.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
```

The structure of each of these tables is described in detail at: https://ohdsi.github.io/CommonDataModel/cdm53.html#Vocabulary_Tables 

## Generating a candidate codelist

For this example we are going to generate a candidate codelist for dementia. Moreover, we're only going to look for codes in the condition domain.

```{r packages,  message=FALSE, warning=FALSE }
library(CodelistGenerator)
library(dplyr)
library(DT)
```

First, let's do a simple search for a single keyword of "dementia" without any exclusions, without searching synonyms, and including descendents but not ancestors.
```{r,  message=FALSE, warning=FALSE, results='hide'  }
dementia_codes1<-get_candidate_codes(keywords="dementia",
                    domains="Condition",
                    search.synonyms = FALSE,
                    fuzzy.match = FALSE,
                    exclude = NULL,
                    include.descendants = TRUE,
                    include.ancestor = FALSE,
                    concept=concept,
                    concept_ancestor = concept_ancestor,
                    concept_synonym = concept_synonym)
```

```{r,  message=FALSE, warning=FALSE }
datatable(dementia_codes1,
          options = list(
  searching = FALSE,
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```


Do we pick up any more codes if we add alzheimer as a key word?
```{r,  message=FALSE, warning=FALSE, results='hide'  }
dementia_codes2<-get_candidate_codes(keywords=c("dementia","alzheimer"),
                    domains="Condition",
                    search.synonyms = FALSE,
                    fuzzy.match = FALSE,
                    exclude = NULL,
                    include.descendants = TRUE,
                    include.ancestor = FALSE,
                    concept=concept,
                    concept_ancestor = concept_ancestor,
                    concept_synonym = concept_synonym) 
```

What extra codes do we pick up?
```{r,  message=FALSE, warning=FALSE }
new_codes_1_2<-bind_rows(dementia_codes1, dementia_codes2) %>% 
  group_by(concept_id) %>% 
  mutate(seq=length(concept_name)) %>%
  filter(seq==1)

datatable(new_codes_1_2,
          options = list(
  searching = FALSE,
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```


Now, we could search synonyms and see if we pick up any extra codes
```{r,  message=FALSE, warning=FALSE, results='hide'  }
dementia_codes3<-get_candidate_codes(keywords=c("dementia","alzheimer"),
                    domains="Condition",
                    search.synonyms = TRUE,
                    fuzzy.match = FALSE,
                    exclude = NULL,
                    include.descendants = TRUE,
                    include.ancestor = FALSE,
                    concept=concept,
                    concept_ancestor = concept_ancestor,
                    concept_synonym = concept_synonym)
```

What extra codes do we pick up from searching synonyms?
```{r,  message=FALSE, warning=FALSE }
new_codes_2_3<-bind_rows(dementia_codes2, dementia_codes3) %>% 
  group_by(concept_id) %>% 
  mutate(seq=length(concept_name)) %>%
  filter(seq==1)

datatable(new_codes_2_3,
          options = list(
  searching = FALSE,
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

Ok, now let' see what we get from also doing a fuzzy search
```{r,  message=FALSE, warning=FALSE, results='hide'  }
dementia_codes4<-get_candidate_codes(keywords=c("dementia","alzheimer"),
                    domains="Condition",
                    search.synonyms = TRUE,
                    fuzzy.match = TRUE,
                    exclude = NULL,
                    include.descendants = TRUE,
                    include.ancestor = FALSE,
                    concept=concept,
                    concept_ancestor = concept_ancestor,
                    concept_synonym = concept_synonym)
```

What extra codes do we pick up?
```{r,  message=FALSE, warning=FALSE }
new_codes_3_4<-bind_rows(dementia_codes3, dementia_codes4) %>% 
  group_by(concept_id) %>% 
  mutate(seq=length(concept_name)) %>%
  filter(seq==1)

datatable(new_codes_3_4,
          options = list(
  searching = FALSE,
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

And this is what the candidate codelist now looks like
```{r,  message=FALSE, warning=FALSE }
datatable(dementia_codes4,
          options = list(
  searching = FALSE,
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```


## Review mappings from source vocabularies
Perhaps we want to see what ICD10CM codes map to our candidate codelist. We can get these by running
```{r,  message=FALSE, warning=FALSE }
icd_mappings<-show_mappings(candidate_codelist=dementia_codes4,
source_vocabularies="ICD10CM",
concept=concept,
concept_relationship=concept_relationship)

datatable(icd_mappings,
          options = list(
  searching = FALSE,
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

```{r,  message=FALSE, warning=FALSE }
read_mappings<-show_mappings(candidate_codelist=dementia_codes4,
source_vocabularies="Read",
concept=concept,
concept_relationship=concept_relationship)

datatable(read_mappings,
          options = list(
  searching = FALSE,
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```
